Hypertension and microbiome

### 2019-05-06 articletwo.Rmd
Scatter plot

    if("clean" %in% names(args)) {
        unlink("rds", recursive=TRUE)
        unlink("cache", recursive=TRUE)
    }
    if ("tags" %in% names(args)) 
        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',
              ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
}
```

@@

```{r maaslin runs, results = 'hide'} 
maaslin <- list()
maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
                forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03",
                           "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
                taxa = "Halapricum.Archaea-")
```

@@

scatter.meta <- meta(pseq.genus) %>%
    tibble::rownames_to_column("id") %>%
      mutate(MAP = oMAP,
           PULSEPRESSURE = oPULSEPRESSURE,
           SYSTM = oSYSTM,
           DIASM = oDIASM) %>%
    gather("keym", "valuem", c(SYSTM, DIASM, MAP, PULSEPRESSURE))

scatter.abund <- abundances(pseq.genus, transform = "compositional") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id") %>%
    setNames(tolower(replace.brackets(names(.)))) %>%
    select(id, scatter.signf.bacteria) %>%
    gather(key, value, -id) %>%
    mutate(value = asin(sqrt(value)))

@@

scatterplot <- ggplot(scatter.full, aes(y = valuem, x = value)) +
    facet_wrap(~key + keym, ncol=4, scales = "fixed") +
    geom_point(size=0.1, shape=1) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y =  -4,
                 parse = TRUE,
                 size = 2.5) +
    ylab("asin(sqrt(abundance))") +
    ggplot2::theme_minimal()

ggsave(file = "cache/scatter.png", plot = scatterplot, height = 20, width = 10, dpi = 300)
```

>>>>

    if("clean" %in% names(args)) {
        unlink("rds", recursive=TRUE)
        unlink("cache", recursive=TRUE)
    }
    if ("tags" %in% names(args)) 
        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',
              ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
    if ("loadlast" %in% names(args) )
        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
}
```

@@

```{r maaslin runs, results = 'hide'} 
maaslin <- list()

genus.core.bacteria <- pseq.genus.core %>%
    abundances %>%
    rownames %>%
    replace.brackets %>%
    mygrep(".Bacteria$", .)

maaslin$genus <- maaslinwrapper(pseq = pseq.genus.core,
                looped = c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"),
                forced = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03",
                           "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
                taxa = paste0(genus.core.bacteria, collapse = ","))
```

@@

scatter.meta <- meta(pseq.genus) %>%
    tibble::rownames_to_column("id") %>%
      mutate(MAP = oMAP,
           PULSEPRESSURE = oPULSEPRESSURE,
           SYSTM = oSYSTM,
           DIASM = oDIASM)

scatter.abund <- abundances(pseq.genus, transform = "clr") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id") %>%
    setNames(tolower(replace.brackets(names(.)))) %>%
    select(id, scatter.signf.bacteria) %>%
    mutate_at(.vars = vars(scatter.signf.bacteria), .funs = funs(myscale(.)))

@@

scatter.combinations <- expand.grid(cc("SYSTM", "DIASM", "PULSEPRESSURE", "MAP"),
                                    c2l(scatter.signf.bacteria)) %>%
    setNames(c("rr", "abund"))

scatter.plot <- function(df,
                         maas.results,
                         rr,
                         abund,
                         covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB",
                                        "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08",
                                        "BL_USE_RX_C09"),
                         prediction = TRUE,
                         plots = TRUE) {
    fo <- sprintf("%s ~ %s + %s", rr, abund, paste0(covariates, collapse="+"))
    model <- lm(as.formula(fo), data = df)

    if (prediction == FALSE) return(model %>% summary)

    df.predict <- data.frame(BPV = seq(-5,5,0.1)) %>%
        mutate(SEX = as.factor(0),
               BL_AGE = df %>% pull(BL_AGE) %>% mean,
               BMI = df %>% pull(BMI) %>% mean,
               CURR_SMOKE = as.factor(0),
               Q57X = as.factor(1),
               DIAB = 0,
               BL_USE_RX_C03 = as.factor(0),
               BL_USE_RX_C07 = as.factor(0),
               BL_USE_RX_C08 = as.factor(0),
               BL_USE_RX_C09 = as.factor(0),
               !!abund := seq(-5,5,0.1)
               )
    prediction <- cbind(df.predict, predict(model, df.predict, interval = 'confidence'))

    if (plots == FALSE) return(prediction)

    feature <- firstup(gsub(".bacteria", "", replace.brackets(abund)))
    qval.maaslin <- maas.results %>%
        dplyr::filter(Feature == feature, Value == rr) %>%
        nrow

    if (qval.maaslin == 0) { return (ggplot() + ggplot2::theme_minimal()) }
    pval <- pub.p(summary(model)$coef[2,4])
    rsqr <- signif(summary(model)$adj.r.squared, 2)
    slope <- signif(model$coef[[2]], 2)

    ggplot(prediction, aes_string(x = abund, y = "fit")) +
        geom_line() +
        geom_point(size=0.1, shape=1, aes_string(x = abund, y = rr, color = "SEX"),
                   data = df) +
        labs(title = sprintf("Adj R2 = %s, Slope = %s, p = %s", rsqr, slope, pval)) +
        xlab(feature) +
        ylab(rr) +
        ylim(0,200) +
        ggplot2::theme_minimal() +
        theme(plot.title = element_text(size=8)) +
        theme(legend.position="none")
}

scatterplots <- lapply(data.frame(t(scatter.combinations)),
                       function(x)
                           scatter.plot(scatter.full, maas.results.df.signf, x$rr, x$abund))

scatterplot <- do.call("grid.arrange", c(scatterplots, ncol=4))
ggsave(file = "cache/scatter.png", plot = scatterplot, height = 20, width = 10, dpi = 300)
```
